---
title: "HVPG TRT and Power Analysis"
always_allow_html: true
output:
  html_document:
    df_print: paged
    keep_md: yes
    toc: yes
    toc_float:
      toc_collapsed: yes
  pdf_document:
    toc: yes
    keep_tex: yes
---

# Aims

Here, I will analyse the HVPG dataset, evaluate aspects of its test-retest reliability, and perform a power analysis for comparing treatments.

# Libraries

```{r message=FALSE, warning=TRUE}
library(tidyverse)
library(relfeas)
library(readxl)
library(hrbrthemes)
library(extrafont)
library(lme4)
library(lmerTest)
library(effsize)
library(broom)
library(RColorBrewer)
library(knitr)
library(cowplot)
library(progress)
library(kableExtra)
library(perm)
library(ggbeeswarm)
library(psych)
library(permuco)
library(metafor)
library(viridis)
library(pwr)

extrafont::loadfonts(quiet = T)

theme_set(theme_ipsum_rc())
nsims <- 1e4
overwrite <- FALSE

knitr::opts_chunk$set(fig.path = "figures/", dev="png", dpi=600,
                      warning=FALSE, message=FALSE)

set.seed(42)
```


# Data

```{r}
trt_tidy <- read_excel("../RawData/TEST_RETEST_FINA_DATABASE_TIDY.xlsx") %>% 
  mutate(author = str_match(Description, "(^\\w*)")[,2]) %>% 
  select(-contains("≦"))

trt_wide <- read_excel("../RawData/TEST_RETEST_FINA_DATABASE.xlsx") %>% 
  select(New_Description, `Serial number`) %>% 
  select(-contains("≦"))

trt_studydemog <- read_excel("../RawData/FINAL_AGGREGATED_ICCs-2.xlsx") %>% 
  select(Study, Perc_Alc, Perc_Decomp, Days=TIMEDAYS, 
         Centre = CENTER,
         n_Patients = NUMBEROFPATIENTs) %>% 
  mutate(Centre = ifelse(Centre == 1, yes = "Multi-centre", "Single-centre"))
```

Now let's add the new new description to the trt_tidy sheet

```{r}
trt_tidy <- trt_tidy %>% 
  left_join(trt_wide)
```


# Study Names

```{r}
studynames <- trt_tidy %>% 
  select(Study = New_Description, 
         Technique = `Technique - Balloon/ catheter`) %>% 
  unique() %>% 
  mutate(Technique = ifelse(Technique=="Wedged Catheter",
                            yes = "Wedged Catheter",
                            no = "Balloon-tipped Catheter")) %>% 
  mutate(Technique = ifelse(is.na(Technique),
                            yes = "Balloon-tipped Catheter",
                            no = Technique)) %>% 
  mutate(Catheter = ifelse(Technique=="Wedged Catheter",
                           yes="Wedged", no="Balloon tip"))
```


# Test-Retest Analysis


## As if one study

Now, we look at the data as if it were all one study, however we also divide by whether the study contains decompensated patients.

```{r}
trt_all <- trt_tidy %>% 
  filter(Description != "Spahr. Octreotide") %>%
  select(-Study) %>%
  rename(Study= New_Description) %>% 
  left_join(trt_studydemog) %>% 
  mutate(decomp = ifelse(Perc_Decomp >0 , 
                           yes="Includes Decompensated",
                           no = "Only Compensated")) %>%
  group_by(decomp) %>% 
  nest() %>% 
  mutate(trt = map(data, ~relfeas::trt(data = .x, 
                                       values='PP', 
                                       cases = "Serial number", 
                                       rater = 'MEASUREMENT' )$tidy)) %>% 
  select(-data) %>% 
  unnest(trt)
  

trt_all

saveRDS(trt_all, "../Cluster/trt_all.rds")

kable(trt_all, digits=2)
```

I'll also create a tidy version of this to complement the individual studies.

```{r}
trt_all_tidy <- trt_all %>% 
  mutate(signvar_sd = signvar_sd * mean) %>% 
  ungroup() %>% 
  select(Patients = decomp, 
         Mean = mean, CV = cv, 
         WSCV = wscv, ICC = icc,
         SDD = sdd,
         "Change SD" = signvar_sd) %>% 
  arrange(desc(Patients))

kable(trt_all_tidy, digits=2)
```



So, overall, we estimate that our smallest detectable difference in an individual is `r round(trt_all$sdd, 1)[2]`mmHg for only compensated patients, and `r round(trt_all$sdd, 1)[1]`mmHg for studies including decompensated patients

```{r}
trt_all_detailed <- trt_tidy %>% 
  filter(Description != "Spahr. Octreotide") %>%
  select(-Study) %>%
  rename(Study= New_Description) %>% 
  left_join(trt_studydemog) %>% 
  mutate(decomp = ifelse(Perc_Decomp >0 , 
                           yes="Includes Decompensated",
                           no = "Only Compensated")) %>%
  filter(Description != "Spahr. Octreotide") %>% 
  group_by(decomp) %>% 
  nest() %>% 
  mutate(trt = map(data, ~relfeas::trt(data = .x, 
                                       values='PP', 
                                       cases = "Serial number", 
                                       rater = 'MEASUREMENT' )))

trt_all_detailed$decomp[1]
trt_all_detailed$trt[[1]]$sdd

trt_all$sdd_l <- trt_all_detailed$trt[[1]]$sdd$lbound
trt_all$sdd_u <- trt_all_detailed$trt[[1]]$sdd$ubound

trt_all_detailed$decomp[2]
trt_all_detailed$trt[[2]]$sdd

trt_all$sdd_l[2] <- trt_all_detailed$trt[[2]]$sdd$lbound
trt_all$sdd_u[2] <- trt_all_detailed$trt[[2]]$sdd$ubound

trt_all_detailed$decomp[1]
trt_all_detailed$trt[[1]]$sddm

trt_all$sddm   <- trt_all_detailed$trt[[1]]$sddm$value*100
trt_all$sddm_l <- trt_all_detailed$trt[[1]]$sddm$lbound*100
trt_all$sddm_u <- trt_all_detailed$trt[[1]]$sddm$ubound*100

trt_all_detailed$decomp[2]
trt_all_detailed$trt[[2]]$sddm

trt_all$sddm[2]   <- trt_all_detailed$trt[[2]]$sddm$value*100
trt_all$sddm_l[2] <- trt_all_detailed$trt[[2]]$sddm$lbound*100
trt_all$sddm_u[2] <- trt_all_detailed$trt[[2]]$sddm$ubound*100

```


Now, let's prepare this as a table for below the forest plot

```{r}
overall_n <- map_dbl(trt_all_detailed$data, nrow)

overall <- trt_all %>% 
  ungroup() %>% 
  rename(Study = decomp) %>% 
  mutate(decomp = "Overall",
         Catheter = "Balloon tip",
         n = overall_n) %>% 
  arrange(desc(Study)) %>% 
  mutate(Study = ifelse(Study=="Includes Decompensated",
                        "Includes Decompensated*",
                        Study))

overall
```



## Grouped by study

```{r}
trt_study <- trt_tidy %>% 
  group_by(New_Description) %>% 
  nest() %>% 
  mutate(outcomes = map(data, ~ trt( data=.x, values='PP', 
                                     cases = "Serial number", rater = 'MEASUREMENT' )))

trt_study <- trt_study %>% 
  mutate(n = map_dbl(data, nrow))
  

tidytrt <- map_df(trt_study$outcomes, 'tidy') %>% 
  mutate(n = trt_study$n) %>% 
  mutate(Study = trt_study$New_Description) %>% 
  select(Study, everything()) %>% 
  left_join(studynames) %>% 
  left_join(trt_studydemog) %>% 
  mutate(decomp = ifelse(Perc_Decomp >0 , 
                           yes="Includes Decompensated",
                           no = "Only Compensated")) %>% 
  mutate(decomp = fct_inorder(decomp)) %>% 
  select(-n_Patients)


knitr::kable(tidytrt, digits = 2)
```


Now let's make a table for the paper with the relevant things.

```{r}
tidytrt_table <- tidytrt %>% 
  mutate(signvar_sd = signvar_sd * mean) %>% 
  select(decomp, 
         Study,
         n,
         "Decompensated (%)" = Perc_Decomp,
         "Alcoholic (%)" = Perc_Alc,
         "Mean Days Elapsed" = Days,
         Mean = mean, CV = cv, 
         WSCV = wscv, ICC = icc,
         SDD = sdd,
         "Change SD" = signvar_sd) %>% 
  arrange(desc(decomp), Study)

decomp_change <- head(
  which(
    tidytrt_table$decomp == tail(tidytrt_table$decomp, 1)), 1 )

tidytrt_kable <- knitr::kable(tidytrt_table[,-1], digits=2) %>% 
  kable_styling("striped", full_width = F) %>%
  pack_rows(head(tidytrt_table$decomp, 1), 1, decomp_change-1) %>%
  pack_rows(tail(tidytrt_table$decomp, 1), decomp_change, nrow(tidytrt_table))

tidytrt_kable

# save_kable(tidytrt_kable, file = "figures/tidytrt_kable.jpg")
```



### ICC 

Let's have a look at the ICC as a measure of between-subject differentiability (i.e. reliability).





```{r icc_forest, fig.height=9, fig.width=8}
icc_out <- select(tidytrt, Study, icc, icc_l, icc_u, decomp, n) %>% 
  left_join(studynames) %>% 
  mutate(decomp = factor(decomp, levels=c(
    "Includes Decompensated", "Only Compensated"))) %>% 
  arrange(decomp, icc) %>% 
  bind_rows(overall) %>% 
  mutate(Study = fct_inorder(Study))

ICCs <- ggplot(icc_out,aes(x=icc,y=Study, 
                           colour=Catheter)) +
  #geom_rect(aes(xmin=0.8, xmax=1, ymin=-Inf, ymax=Inf),
  #          alpha = .1, fill="grey", colour="grey") +
  facet_grid(decomp~., scales="free", space="free") +
  geom_point(aes(size=log(n)), shape=18) + 
  scale_x_continuous(breaks = seq(0, 1, by=0.2))+
  geom_errorbarh(aes(xmax = icc_u, xmin = icc_l), height = 0.2) +
  #geom_vline(xintercept = 1, linetype = "longdash") +
  labs(y="", x="Intraclass Correlation Coefficient (ICC)") +
  theme(text = element_text(size=20)) +
  ggtitle("ICC (95% CI)") +
  scale_colour_manual(values = c("black", "red")) +
  coord_cartesian(xlim = c(-0.1, 1)) +
                  #ylim = c(0,12)) +
  guides(size = FALSE) +
  NULL

ICCs + theme(legend.position = "none")
```

This analysis constitutes a mega-analysis: we have all the original data and the overall estimates are performed using the original data estimates. However, in order to asess the study heterogeneity, I will run a classical meta-analysis. I will perform a Fisher's z transformation on the ICC values, perform a classic meta-analysis, and assess the heterogeneity.


```{r}
dat <- icc_out %>% 
  filter(Study != "Spahr 2007") %>% 
  filter(decomp != "Overall") %>% 
  filter(n > 4) %>% 
  escalc(measure="ZCOR", ri=icc, ni=n, data=., slab=Study) 

res <- rma(yi, vi, data=dat) 
res 
```

### SDD

#### Absolute

Let's have a look at the study-by-study SDD, but in raw units

```{r sdd_absolute_forest, fig.height=9, fig.width=8}

sdd_out <- map_df(trt_study$outcomes, "sdd") %>% 
  mutate(Study = trt_study$New_Description) %>% 
  rename(sdd=value, sdd_l = lbound, sdd_u = ubound) %>% 
  left_join(studynames) %>% 
  left_join(select(tidytrt, Study, decomp, n)) %>% 
  left_join(trt_studydemog) %>% 
  mutate(decomp = ifelse(Perc_Decomp >0 , 
                           yes="Includes Decompensated",
                           no = "Only Compensated")) %>% 
  mutate(decomp = factor(decomp, levels=c(
    "Includes Decompensated", "Only Compensated"))) %>% 
  arrange(decomp, desc(sdd)) %>% 
  bind_rows(overall) %>% 
  mutate(Study = fct_inorder(Study))

SDDs <- ggplot(sdd_out,aes(x=sdd,y=Study, 
                           colour=Catheter)) +
  facet_grid(decomp~., scales="free", space="free") +
  geom_point(aes(size=log(n)), shape=18) + 
  scale_x_continuous(breaks = seq(0, 20, by=2))+
  geom_errorbarh(aes(xmax = sdd_u, xmin = sdd_l), height = 0.15) +
  #geom_vline(xintercept = 1, linetype = "longdash") +
  labs(y="", x="Smallest Detectable Difference (mmHg)") +
  theme(text = element_text(size=20)) +
  ggtitle("SDD (95% CI)") +
  scale_colour_manual(values = c("black", "red")) +
  guides(colour=FALSE, shape=FALSE) +
  #scale_shape_manual(values = c(18, 19)) +
  coord_cartesian(xlim=c(0,16)) +
  guides(size = FALSE) +
  NULL

SDDs  #+ annotate("rect", xmin = 0.75, xmax = 1, ymin="Spahr 2007", ymax="Blei 1987", alpha = .2, fill="grey")
```

And here we run the meta-analysis again to assess heterogeneity.  We can't use the SDD because it has asymmetric confidence intervals, and it cannot be Fisher's z transformed.  So we can run a meta-analysis based on the average absolute variation to test the heterogeneity.


```{r}
dat <- trt_study %>% 
  mutate(
    mean_abs_var = map_dbl(outcomes, ~mean(sqrt(.x$absvars))),
    se_abs_var = map_dbl(outcomes, ~sd(sqrt(.x$absvars)) / sqrt(length(.x$absvars)))
  ) %>% 
  filter(New_Description != "Spahr 2007") %>% 
  ungroup() %>% 
  mutate(
    yi = mean_abs_var,
    vi = se_abs_var^2) %>% 
  select(-data, -outcomes) %>% 
  escalc(measure="MN", yi=yi, vi=vi, ni=n, data=., slab=New_Description) 

transf.sqrt <- function (xi, ...) 
{
    zi <- sqrt(xi)
    zi[xi < 0] <- 0
    return(c(zi))
}

res <- rma(yi, vi, data=dat)
res 
```


#### Percentage

Now the study-by-study SDD, but in percentages

```{r sdd_perc_forest, fig.height=9, fig.width=8}

sddm_out <- map_df(trt_study$outcomes, "sddm") %>% 
  mutate(Study = trt_study$New_Description) %>% 
  rename(sddm=value, sddm_l = lbound, sddm_u = ubound) %>% 
  mutate(sddm=100*sddm, sddm_l = 100*sddm_l, sddm_u = 100*sddm_u) %>% 
  left_join(studynames) %>% 
  left_join(select(tidytrt, Study, decomp, n)) %>% 
  left_join(trt_studydemog) %>% 
  mutate(decomp = ifelse(Perc_Decomp >0 , 
                           yes="Includes Decompensated",
                           no = "Only Compensated")) %>% 
  mutate(decomp = factor(decomp, levels=c(
    "Includes Decompensated", "Only Compensated"))) %>% 
  arrange(decomp, desc(sddm)) %>% 
  bind_rows(overall) %>% 
  mutate(Study = fct_inorder(Study))

SDDms <- ggplot(sddm_out,aes(x=sddm,y=Study, 
                           colour=Catheter)) +
  facet_grid(decomp~., scales="free", space="free") +
  geom_point(aes(size=log(n)), shape=18) + 
  scale_x_continuous(breaks = seq(0, 80, by=10))+
  geom_errorbarh(aes(xmax = sddm_u, xmin = sddm_l), height = 0.15) +
  #geom_vline(xintercept = 1, linetype = "longdash") +
  labs(y="", x="Smallest Detectable Difference (%mmHg)") +
  theme(text = element_text(size=20)) +
  ggtitle("%SDD (95% CI)") +
  scale_colour_manual(values = c("black", "red")) +
  guides(colour=FALSE, shape=FALSE) +
  #scale_shape_manual(values = c(18, 19)) +
  coord_cartesian(xlim=c(0,80)) +
  guides(size = FALSE) +
  NULL

SDDms  #+ annotate("rect", xmin = 0.75, xmax = 1, ymin="Spahr 2007", ymax="Blei 1987", alpha = .2, fill="grey")
```

And here we run the meta-analysis to assess heterogeneity.  We can't use the SDD% because it has asymmetric confidence intervals, and it cannot be Fisher's z transformed.  So we can run a meta-analysis based on the average absolute percentage variation to test the heterogeneity.


```{r}
dat <- trt_study %>% 
  mutate(
    mean_abs_var = map_dbl(outcomes, ~mean(sqrt(.x$absvars / .x$means))),
    se_abs_var = map_dbl(outcomes, ~sd(sqrt(.x$absvars / .x$means)) / 
                           sqrt(length(.x$absvars)))
  ) %>% 
  filter(New_Description != "Spahr 2007") %>% 
  ungroup() %>% 
  mutate(
    yi = mean_abs_var,
    vi = se_abs_var^2)

res <- rma(yi, vi, data=dat) 
res 
```


### Figures together

```{r sdd_together_forests, fig.height=6, fig.width=14, message=FALSE, warning=FALSE}
SDD_together <- cowplot::plot_grid(SDDs, SDDms, ncol = 2,
                                   labels = c("B", "C"),
                                   label_size = 30)

SDD_together
```

<!-- ### Even more together -->

<!-- ```{r all_forest, fig.height=12, fig.width=14} -->
<!-- ICCs_2 <- ICCs + coord_cartesian(xlim=c(0,1)) + -->
<!--   scale_shape_discrete("Patients") -->

<!-- plot_grid(ICCs_2, SDD_together,  -->
<!--           ncol=1, labels=c("A", ""),  -->
<!--           label_size=30, rel_heights = c(3,2)) -->
<!-- ``` -->

Alternatively, all in a row

```{r all_forest_row, fig.height=8, fig.width=16, message=FALSE, warning=FALSE}
ICCs_row <- ICCs + guides(colour=FALSE, shape=FALSE)

forest_legend <- get_legend(ICCs + 
                              theme(legend.position="bottom") +
                              scale_shape_discrete("Patients:") +
                              scale_colour_manual("Catheter:",
                                values=c("black", "red")))

forest_row_legendless <- plot_grid(ICCs_row, SDDs, SDDms,
          nrow=1, labels=c("A", "B", "C"), 
          label_size=30)

forest_row <- plot_grid(forest_row_legendless, forest_legend,
                        nrow=2, rel_heights = c(15,1))

forest_row

ggsave(forest_row, filename = "figures/forest_three_row.png", width = 16, height = 8)
```

### Just ICC and SDD

```{r two_forest_row, fig.height=9, fig.width=12, message=FALSE, warning=FALSE}
forest_row2_legendless <- plot_grid(ICCs_row, SDDs,
          nrow=1, labels=c("A", "B"), 
          label_size=30)

forest_row2 <- plot_grid(forest_row2_legendless, forest_legend,
                        nrow=2, rel_heights = c(15,1))

forest_row2
```



## Are they different

To test the difference between the groups, we'll bootstrap the difference.  That means we take 1000 random samples from each group of the same size with replacement, and calculate the difference to get the null distribution. Then we compare the difference we see to the bootstrap distribution.


```{r}
trt_compare <- trt_tidy %>% 
  filter(Description != "Spahr. Octreotide") %>%
  select(-Study) %>%
  rename(Study= New_Description) %>% 
  left_join(trt_studydemog) %>% 
  mutate(decomp = ifelse(Perc_Decomp >0 , 
                           yes="Includes Decompensated",
                           no = "Only Compensated")) %>%
  group_by(decomp) %>% 
  nest()
```

### ICC

```{r}
bootstrap_icc_single <- function(data) {
  
  sample_ids <- unique(data$`Serial number`)
  ids <- sample(sample_ids, length(sample_ids), replace=TRUE)
  
  data_nested <- data %>% 
    select(`Serial number`, MEASUREMENT, PP) %>% 
    nest_by(`Serial number`)
  
  boot_sample <- tibble(
    `Serial number` = ids
  ) %>% 
    left_join(data_nested, by="Serial number") %>% 
    select(-`Serial number`) %>% 
    mutate(boot_serial = 1:n()) %>% 
    unnest(data) %>% 
    relfeas::trt_widify("PP", "boot_serial", "MEASUREMENT")

  boot_icc <- suppressMessages(
    suppressWarnings(
      psych::ICC(boot_sample[,c(2,3)])$results$ICC[2] ) )
  return(boot_icc)
}

bootstrap_icc <- function(n, data, colname=NULL) {
  
  out <- tibble(
    n = 1:n
  ) %>% 
    mutate(boot_icc = map_dbl(n, ~bootstrap_icc_single(data)))
  
  if(!is.null(colname)) {
    colnames(out)[2] <- colname
  }
  
  return(out)
}
```



```{r}
trt_compare_iccvals <- trt_compare %>% 
  mutate(boot = map2(data, decomp, ~bootstrap_icc(10000, .x))) %>% 
  select(-data) %>% 
  unnest(boot)
```

```{r}
trt_compare_icc <- trt_compare_iccvals %>% 
  spread(decomp, boot_icc) %>% 
  mutate(dif = `Only Compensated` - `Includes Decompensated`)

quantile(trt_compare_icc$dif, c(0.025, 0.5, 0.975)) # 95% CI
1-sum(trt_compare_icc$dif > 0)/10000 # one-sided p value
```

Not significant

### SDD


```{r}
bootstrap_sdd_single <- function(data) {
  
  sample_ids <- unique(data$`Serial number`)
  ids <- sample(sample_ids, length(sample_ids), replace=TRUE)
  
  data_nested <- data %>% 
    select(`Serial number`, MEASUREMENT, PP) %>% 
    nest_by(`Serial number`)
  
  boot_sample <- tibble(
    `Serial number` = ids
  ) %>% 
    left_join(data_nested, by="Serial number") %>% 
    select(-`Serial number`) %>% 
    mutate(boot_serial = 1:n()) %>% 
    unnest(data) %>% 
    relfeas::trt_widify("PP", "boot_serial", "MEASUREMENT")

  boot_sdd <- suppressMessages(
    suppressWarnings(
      agRee::agree.sdd(as.matrix(boot_sample[,c(2,3)]))$value ) )
  return(boot_sdd)
}


bootstrap_sdd_single <- function(data) {
  
  sample_ids <- unique(data$`Serial number`)
  ids <- sample(sample_ids, length(sample_ids), replace=TRUE)
  
  data_nested <- data %>% 
    select(`Serial number`, MEASUREMENT, PP) %>% 
    nest_by(`Serial number`)
  
  boot_sample <- tibble(
    `Serial number` = ids
  ) %>% 
    left_join(data_nested, by="Serial number") %>% 
    select(-`Serial number`) %>% 
    mutate(boot_serial = 1:n()) %>% 
    unnest(data) %>% 
    relfeas::trt_widify("PP", "boot_serial", "MEASUREMENT")

  boot_sdd <- suppressMessages(
    suppressWarnings(
      agRee::agree.sdd(as.matrix(boot_sample[,c(2,3)]))$value ) )
  return(boot_sdd)
}

bootstrap_sdd <- function(n, data, colname=NULL) {
  
  out <- tibble(
    n = 1:n
  ) %>% 
    mutate(boot_sdd = map_dbl(n, ~bootstrap_sdd_single(data)))
  
  if(!is.null(colname)) {
    colnames(out)[2] <- colname
  }
  
  return(out)
}

# bootstrap_sdd_single(trt_compare$data[[1]])
# bootstrap_sdd(100, trt_compare$data[[1]], "test")
```

And now we test

```{r}
trt_compare_sddvals <- trt_compare %>% 
  mutate(boot = map2(data, decomp, ~bootstrap_sdd(10000, .x))) %>% 
  select(-data) %>% 
  unnest(boot)
```

```{r}
trt_compare_sdd <- trt_compare_sddvals %>% 
  spread(decomp, boot_sdd) %>% 
  mutate(dif = `Includes Decompensated` - `Only Compensated`)

quantile(trt_compare_sdd$dif, c(0.025, 0.5, 0.975)) # 95% CI
1-sum(trt_compare_sdd$dif > 0)/10000 # one-sided p value
```

### SDDM

Percentage SDD


```{r}
bootstrap_sddm_single <- function(data) {
  
  sample_ids <- unique(data$`Serial number`)
  ids <- sample(sample_ids, length(sample_ids), replace=TRUE)
  
  data_nested <- data %>% 
    select(`Serial number`, MEASUREMENT, PP) %>% 
    nest_by(`Serial number`)
  
  boot_sample <- tibble(
    `Serial number` = ids
  ) %>% 
    left_join(data_nested, by="Serial number") %>% 
    select(-`Serial number`) %>% 
    mutate(boot_serial = 1:n()) %>% 
    unnest(data) %>% 
    relfeas::trt_widify("PP", "boot_serial", "MEASUREMENT")

  boot_sddm <- suppressMessages(
    suppressWarnings(
      agRee::agree.sddm(as.matrix(boot_sample[,c(2,3)]))$value ) )
  return(boot_sddm)
}

bootstrap_sddm <- function(n, data, colname=NULL) {
  
  out <- tibble(
    n = 1:n
  ) %>% 
    mutate(boot_sddm = map_dbl(n, ~bootstrap_sddm_single(data)))
  
  if(!is.null(colname)) {
    colnames(out)[2] <- colname
  }
  
  return(out)
}

# bootstrap_sddm_single(trt_compare$data[[1]])
# bootstrap_sddm(100, trt_compare$data[[1]], "test")
```

And now we test

```{r}
trt_compare_sddmvals <- trt_compare %>% 
  mutate(boot = map2(data, decomp, ~bootstrap_sddm(10000, .x))) %>% 
  select(-data) %>% 
  unnest(boot)
```

```{r}
trt_compare_sddm <- trt_compare_sddmvals %>% 
  spread(decomp, boot_sddm) %>% 
  mutate(dif = `Includes Decompensated` - `Only Compensated`)

quantile(trt_compare_sddm$dif, c(0.025, 0.5, 0.975)) # 95% CI
1-sum(trt_compare_sddm$dif > 0)/10000 # one-sided p value
```





# Causes of differences

Here, we perform an exploratory analysis of which factors may contribute to differences

```{r, warning=FALSE}
trt_wide <- trt_tidy %>% 
  select(-Study) %>%
  rename(Study= New_Description) %>% 
  spread(MEASUREMENT, PP) %>% 
  rename(Meas1 = `1`,
         Meas2 = `2`) %>% 
  mutate(change = Meas2 - Meas1,
         abschange=abs(change),
         meanval = (Meas1 + Meas2)/2) %>% 
  left_join(studynames) %>% 
  left_join(trt_studydemog) %>% 
  mutate(Description = as.factor(Description)) %>% 
  filter(Description != "Spahr. Octreotide") %>% 
  mutate(decomp = ifelse(Perc_Decomp == 0,
                         "Only Compensated",
                         "Includes Decompensated"))


trt_wide_c <- trt_wide %>% 
  filter(Perc_Decomp == 0)

trt_wide_dc <- trt_wide %>% 
  filter(Perc_Decomp > 0)
```

## Skewness

### Absolute

Let's just visualise our distribution.

```{r, fig.height=3, fig.width=4}
ggplot(trt_wide, aes(x=abschange)) +
  geom_density(fill="black",alpha=0.5)
```
And between groups

```{r, fig.height=6, fig.width=4}
ggplot(trt_wide, aes(x=abschange)) +
  geom_density(aes(colour=decomp, fill=decomp),alpha=0.5) +
  facet_wrap(decomp~., nrow=2) +
  theme(legend.position="bottom")
```

And let's get some values for that too

```{r}
psych::describe(trt_wide$abschange)
psych::describeBy(trt_wide$abschange, group=trt_wide$decomp)
```


### Signed

Let's just visualise our overall distribution.

```{r, fig.height=3, fig.width=4}
ggplot(trt_wide, aes(x=change)) +
  geom_density(fill="black",alpha=0.5)
```
... and by group

```{r, fig.height=6, fig.width=4}
ggplot(trt_wide, aes(x=change)) +
  geom_density(aes(colour=decomp, fill=decomp),alpha=0.5) +
  facet_wrap(decomp~., nrow=2) +
  theme(legend.position="bottom")
```

For signed differences, let's assess whether they differ from zero.

```{r}
summary(lmer(change ~ 1 + (1|Study), data=trt_wide))

summary(lmer(change ~ 1 + (1|Study), data=trt_wide_c))

summary(lmer(change ~ 1 + (1|Study), data=trt_wide_dc))
```

Not significantly different from zero in either the combined sample or each group.

And some values

```{r}
psych::describe(trt_wide$change)
psych::describeBy(trt_wide$change, group=trt_wide$decomp)
```


### Figures

Let's make some combined figures for the paper

```{r, fig.height=4, fig.width=6}
sign_change_distr <- trt_wide %>% 
  mutate(decomp = as.factor(decomp)) %>% 
  ggplot(aes(x=change, colour=decomp, fill=decomp, group=decomp)) +
  geom_density(aes(colour=decomp, fill=decomp),alpha=0.5) +
  theme(legend.position=c(0.3, 0.9)) +
  scale_color_brewer("Patients", type = "qual", palette = 1) +
  scale_fill_brewer("Patients", type = "qual", palette = 1) +
  guides(fill=FALSE) +
  labs(y = "",
       x = "Signed Change from first to second measurement (mmHg)") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y=element_blank())

sign_change_distr

abs_change_distr <- trt_wide %>% 
  mutate(decomp = as.factor(decomp)) %>% 
  ggplot(aes(x=abschange, colour=decomp, fill=decomp, group=decomp)) +
  geom_density(aes(colour=decomp, fill=decomp),alpha=0.5) +
  theme(legend.position=c(0.5, 0.9)) +
  scale_color_brewer("Patients", type = "qual", palette = 1) +
  scale_fill_brewer("Patients", type = "qual", palette = 1) +
  guides(fill=FALSE) +
  labs(y = "",
       x = "Absolute Change from first to second measurement (mmHg)") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y=element_blank())

abs_change_distr
```









## Percentage Decomp

```{r, warning=FALSE}
permTREND(formula=abschange ~ Perc_Decomp, data=trt_wide,
          method="exact.mc")

permuco::lmperm(abschange ~ Perc_Decomp, data=trt_wide)
```


```{r, fig.height=4, fig.width=8}
decomp_plot <- ggplot(trt_wide, aes(x=Perc_Decomp, y=abschange)) +
  geom_beeswarm(aes(colour=as.factor(Study), 
                    group=as.factor(Perc_Decomp)),
                alpha=0.4, cex=1) +
  guides(colour=FALSE) + 
  geom_smooth(method="lm", se = FALSE) +
  labs(x="Decompensated Patients (%)",
       y="Absolute Change (mmHg)")

decomp_plot
```


## Days elapsed

```{r, warning=FALSE, fig.width=6, fig.height=8}
permuco::lmperm(abschange ~ Days + Perc_Decomp, data=trt_wide)
```

Now let's plot, after correcting for the effect of the patient groups.

```{r, fig.height=4, fig.width=6}
correct_for_decomp <- function(formula) {
  
  formula <- as.formula(formula)
  
  coefficients <- coef(permuco::lmperm(formula, 
                                       data=trt_wide))

  predicted <- as.character(formula[2])
  
  after_decomp_corr <- trt_wide[[predicted]] - 
    trt_wide[["Perc_Decomp"]] * coefficients[which(names(coefficients)=="Perc_Decomp")]
  
  return(after_decomp_corr)
  
}

days_plot <- trt_wide %>% 
  mutate(abschange_dccorr = correct_for_decomp(abschange ~ Days + Perc_Decomp)) %>% 
  ggplot(aes(x=Days, y=abschange_dccorr)) +
  geom_beeswarm(aes(colour=Study, group=Study), alpha=0.4, cex=1) +
  guides(colour=FALSE) + 
  geom_smooth(method="lm", se=FALSE) +
  labs(y="Absolute Change (mmHg)\n(after correction for Decompensated Percentage)",
       x="Days Elapsed between Measurements")

days_plot
```

## Percentage Alcoholic

```{r, warning=FALSE}
permuco::lmperm(abschange ~ Perc_Alc + Perc_Decomp, data=trt_wide)

permTREND(formula=abschange ~ Perc_Alc, data=trt_wide_c,
          method="exact.mc")

permTREND(formula=abschange ~ Perc_Alc, data=trt_wide_dc,
          method="exact.mc")
```

```{r, fig.height=4, fig.width=6}
perc_alc_plot <- trt_wide %>% 
  mutate(abschange_dccorr = correct_for_decomp(abschange ~ Perc_Alc + Perc_Decomp)) %>% 
  ggplot(aes(x=Perc_Alc, y=abschange_dccorr)) +
  geom_beeswarm(aes(colour=Study, group=Study), alpha=0.4, cex=1) +
  guides(colour=FALSE) + 
  geom_smooth(method="lm", se=FALSE) +
  labs(y="Absolute Change (mmHg)\n(after correction for Decompensated Percentage)",
       x="Percentage of Alcoholic Patients (%)")

perc_alc_plot
```

## Multicentre

```{r, warning=FALSE}
permuco::lmperm(abschange ~ Centre + Perc_Decomp, data=trt_wide)
```

```{r, fig.height=4, fig.width=6}
centre_plot <- trt_wide %>% 
  mutate(abschange_dccorr = correct_for_decomp(abschange ~ Centre + Perc_Decomp)) %>% 
  ggplot(aes(x=Centre, y=abschange_dccorr)) +
  geom_beeswarm(aes(colour=Study, group=Study), alpha=0.4, cex=1) +
  guides(colour=FALSE) + 
  geom_smooth(method="lm", se=FALSE) +
  labs(y="Absolute Change (mmHg)\n(after correction for Decompensated Percentage)")

centre_plot
```


## Combined Model

```{r}
permuco::lmperm(abschange ~ Perc_Decomp + Centre + Perc_Alc, data=trt_wide)

permuco::lmperm(abschange ~ Centre + Perc_Alc, data=trt_wide_c)
permuco::lmperm(abschange ~ Centre + Perc_Alc, data=trt_wide_dc)
```
Let's examine why this might be.

```{r}
trt_wide %>% 
  ggplot(aes(y=Perc_Decomp, x=Centre)) + 
  geom_violin()

permTS(formula=Perc_Decomp ~ as.factor(Centre), data=trt_wide,
          method="exact.mc")
```

Most of the single-centre studies have high proportions of decompensated patients.

```{r}
trt_wide %>% 
  ggplot(aes(y=Perc_Alc, x=Centre)) + 
  geom_violin()

permTS(formula=Perc_Alc ~ as.factor(Centre), data=trt_wide,
          method="exact.mc")
```
Similarly, most of the single-centre studies have high proportions of alcoholic patients.

### Days

Reviewers were surprised at the lack of an effect of the number of days elapsed. Perhaps this was obscured by not having included the other confounding factors.

```{r}
permuco::lmperm(abschange ~ Perc_Decomp + Centre + Perc_Alc + Days, data=trt_wide)
```
This does not appear to be the case!


## Mean Value

### Absolute

```{r, warning=FALSE}
permuco::lmperm(abschange ~ meanval + Perc_Decomp, data=trt_wide)

permTREND(formula=abschange ~ meanval, data=trt_wide_c,
          method="exact.mc")

permTREND(formula=abschange ~ meanval, data=trt_wide_dc,
          method="exact.mc")
```

```{r, fig.height=4, fig.width=6}
meanv_abs_plot <- trt_wide %>% 
  mutate(abschange_dccorr = correct_for_decomp(abschange ~ meanval + Perc_Decomp)) %>% 
  ggplot(aes(x=meanval, y=abschange_dccorr)) +
  geom_point(aes(colour=Study, group=Study), alpha=0.4) +
  guides(colour=FALSE) + 
  geom_smooth(method="lm", se=FALSE) +
  labs(y="Absolute Change (mmHg)\n(after correction for Decompensated Percentage)",
       x="Mean Value across Measurements (mmHg)")

meanv_abs_plot
```


### Signed

```{r, warning=FALSE}
meanv_sign <- lmer(change ~ meanval + Perc_Decomp + (1 | Study), data = trt_wide)
summary(meanv_sign)

meanv_sign_nocorr <- lmer(change ~ meanval + (1 | Study), data = trt_wide)
summary(meanv_sign_nocorr)

meanv_sign_c <- lmer(change ~ meanval + (1 | Study), data = trt_wide_c)
summary(meanv_sign_c)

meanv_sign_dc <- lmer(change ~ meanval + (1 | Study), data = trt_wide_dc)
summary(meanv_sign_dc)
```

```{r, fig.height=4, fig.width=6}
correct_for_decomp_lmer <- function(formula) {
  
  formula <- as.formula(formula)
  
  coefficients <- fixef(lmer(formula, data=trt_wide))

  predicted <- as.character(formula[2])
  
  after_decomp_corr <- trt_wide[[predicted]] - 
    trt_wide[["Perc_Decomp"]] * coefficients[which(names(coefficients)=="Perc_Decomp")]
  
  return(after_decomp_corr)
  
}


meanv_signed_plot <- trt_wide %>% 
  mutate(change_dccorr = correct_for_decomp_lmer("change ~ meanval + Perc_Decomp + (1 | Study)")) %>% 
  ggplot(aes(x=meanval, y=change_dccorr)) +
  geom_jitter(aes(colour=Study, group=Study), alpha=0.4, height = 0.2) +
  guides(colour=FALSE) + 
  geom_smooth(method="lm", se=FALSE) +
  labs(y="Signed Change (mmHg)\n(after correction for Decompensated Percentage)",
       x="Mean Value across Measurements (mmHg)")

meanv_signed_plot
```


```{r, fig.height=4, fig.width=6}
meanv_signed_plot_nocorr <- ggplot(data=trt_wide, aes(x=meanval, y=change)) +
  geom_jitter(aes(colour=Study, group=Study), alpha=0.4, height = 0.2) +
  guides(colour=FALSE) + 
  geom_smooth(method="lm", se=FALSE) +
  labs(y="Signed Change (mmHg)\n(jittered slightly to avoid overlap)",
       x="Mean Value across Measurements (mmHg)")

meanv_signed_plot_nocorr
```




## Figures

```{r, fig.height=7, fig.width=9, warning=FALSE}
cowplot::plot_grid(decomp_plot, days_plot,
                   perc_alc_plot, centre_plot, align = "hv", 
                   ncol = 2, labels = "AUTO")
```

```{r, fig.height=7, fig.width=9, warning=FALSE}
cowplot::plot_grid(abs_change_distr, sign_change_distr,
                   meanv_abs_plot, meanv_signed_plot, 
                   align = "hv", 
                   ncol = 2, labels = "AUTO")
```



# Power Analysis for a difference

Now, the core thing we want to do here is to perform a power analysis for examining within-individual effects.

One way of doing this is to use the `signvar_sd` column of the `tidytrt` object. This is the standard deviation of the signed changes, and hence, if we assume a change after an intervention, this is the SD we could imagine being true, and thus, the effect size, the Cohen's Dz, is equal to difference / signvar_sd.  However, this method makes an assumption that everyone changes by exactly the same amount: the effect (before accounting for error) is completely uniform.  This may be the case, but this is the most optimistic scenario.  We should be taking into consideration the possibility of heterogeneous effects.


First, let's make a little plot to show what I mean by homogeneous and heterogeneous effects.

```{r, fig.height=6, fig.width=8}
set.seed(1234)

trt_all_comp <- trt_all[2,]

wscv=trt_all_comp$wscv
meanval=trt_all_comp$mean
cv=trt_all_comp$cv
icc=trt_all_comp$icc

sd_true <- sqrt(icc * (cv * meanval)^2)

n <- 20
delta <- 2

# Homogeneous
cv_delta <- 0

pre_true <- rnorm(n, meanval, sd_true)
pre_meas <- pre_true + rnorm(n, 0, meanval*wscv)

post_true <- pre_true - rnorm(n, delta, cv_delta*delta)
post_meas <- post_true + rnorm(n, 0, meanval*wscv)

hom_true <- tibble::tibble(
  ID = rep(1:n, times=2),
  Outcome = c(pre_true, post_true),
  PrePost = rep(c("Pre", "Post"), each=n),
  Effects = "Homogeneous Effects",
  MeasuredTrue = "True Values"
)

hom_measured <- tibble::tibble(
  ID = rep(1:n, times=2),
  Outcome = c(pre_meas, post_meas),
  PrePost = rep(c("Pre", "Post"), each=n),
  Effects = "Homogeneous Effects",
  MeasuredTrue = "Measured Values"
)

# hom_difference <- tibble::tibble(
#   ID = rep(1:n, times=2),
#   Outcome = c(post_meas-pre_),
#   PrePost = rep(c("Pre", "Post"), each=n),
#   Effects = "Homogeneous",
#   MeasuredTrue = "Difference"
# )


# Heterogeneous
cv_delta <- 0.5

#pre_true <- rnorm(n, meanval, abs(cv*meanval)) # Use same as above
#pre_meas <- pre_true + rnorm(n, 0, abs(meanval*wscv))

post_true <- pre_true - rnorm(n, delta, abs(cv_delta*delta))
post_meas <- post_true + rnorm(n, 0, abs(meanval*wscv))

het_true <- tibble::tibble(
  ID = rep(1:n, times=2),
  Outcome = c(pre_true, post_true),
  PrePost = rep(c("Pre", "Post"), each=n),
  Effects = "Heterogeneous Effects",
  MeasuredTrue = "True Values"
)

het_measured <- tibble::tibble(
  ID = rep(1:n, times=2),
  Outcome = c(pre_meas, post_meas),
  PrePost = rep(c("Pre", "Post"), each=n),
  Effects = "Heterogeneous Effects",
  MeasuredTrue = "Measured Values"
)

# Plot
effects <- bind_rows(hom_true, hom_measured, het_true, het_measured) %>% 
  mutate(MeasuredTrue = fct_inorder(MeasuredTrue),
         Effects = fct_inorder(Effects),
         PrePost = fct_inorder(PrePost),
         ID = as.factor(ID))

ggplot(effects, aes(x=PrePost, y=Outcome, colour=ID, group=ID)) +
  geom_point(size=2) +
  geom_line() +
  facet_grid(Effects~MeasuredTrue) +
  labs(y="Outcome (mmHg)",
       colour="Values",
       x=NULL,
       title="Homogeneous and Heteregeneous Effects",
       subtitle="Homogeneous effects imply that the true underlying change is the same\nin all individuals (hence parallel lines in the true values)") +
  guides(colour=FALSE)

```

So, to summarise, we have underlying true values, and measured values after accounting for measurement error.  The change from before to after the intervention can either be homogeneous (everyone has exactly the same effect), or heterogeneous (effect sizes differ, and some even get harmed by the intervention - about 2.5% as I've chosen the SD of the intervention effect as 50% of the mean effect, so 0 effect is 2 SDs away from the mean effect size, which is approximately 2.5%).  Then, the measured values appear to show more people getting worse after treatment, but this is just due to measurement error.

```{r fig.height=6, fig.width=8}
heterogen_cv <- 0.5

annotations <- tibble(
  x=c(-3, 0),
  text=c(paste0(round(100*pnorm(1/heterogen_cv)), "% experience improvements"),
         paste0(100-round(100*pnorm(1/heterogen_cv)), "% experience worsening")),
  colour = c("#61b096", "#bd7969")
)

heterogen_effects <- tibble(Effect=c(-3, 1)) 

ggplot(heterogen_effects, aes(x=Effect)) +
  geom_area(stat="function", fun = dnorm, fill="#61b096", xlim=c(-3, 0),
            args = list(mean = -1, sd=heterogen_cv), alpha=0.7) +
  geom_area(stat="function", fun = dnorm, fill="#bd7969", xlim=c(0, 1),
            args = list(mean = -1, sd=heterogen_cv), alpha=0.7) +
  annotate("text", x=-2.5, y=0.7, 
           label=paste0(round(100*pnorm(1/heterogen_cv), 1), 
                        "% experience\nimprovements"),
           colour = "#61b096", hjust=0.5) +
  annotate("text", x=0.5, y=0.7, 
           label=paste0(100-round(100*pnorm(1/heterogen_cv), 1), 
                        "% experience\nworsening"),
           colour = "#bd7969", hjust=0.5) +
  annotate("text", x=-0.8, y=0.2, 
           label="Mean\neffect",hjust=0.5) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  geom_vline(xintercept = -1) +
  labs(title="Heterogeneous Effects",
       subtitle=paste0("Using a 50% CV for effect heterogeneity implies that some ",
                       "participants\nmay benefit more and others may even have ",
                       "worsening."))
```

This would imply that for all different sizes of effect, that only 2.3% experience a true worsening. However, when we measure the values, there are also several individuals who will exhibit an apparent worsening (increases from first to second measurement), when their true underlying values exhibited improvements. Let's calculate what fraction of individuals this would be.

## True and Apparent Changes

Here, we can also observe the percentage of individuals who show apparent 10% and 20% changes from baseline.

```{r, eval=FALSE}
apparent_effects <- function(n, delta, cv_delta, wscv=trt_all$wscv, 
                     mean=trt_all$mean, cv=trt_all$cv, icc=trt_all$icc, 
                     decomp) {
  
  wscv <- wscv[decomp]
  mean <- mean[decomp]
  cv   <- cv[decomp]
  
  var <- (cv*mean)^2
  sd_true <- sqrt(var * icc)
  
  
  pre_true <- rnorm(n, mean, sd_true)
  
  pre_meas <- pre_true + rnorm(n, 0, mean*wscv)
  
  post_true <- pre_true - rnorm(n, delta, cv_delta*delta)
  
  post_meas <- post_true + rnorm(n, 0, mean*wscv)
  
  measured <- tibble::tibble(
    ID = rep(1:n, times=2),
    Outcome = c(pre_meas, post_meas),
    PrePost = rep(c("Pre", "Post"), each=n)
  ) %>% 
    spread(PrePost, Outcome)
  
  true <- tibble::tibble(
    ID = rep(1:n, times=2),
    Outcome = c(pre_true, post_true),
    PrePost = rep(c("Pre", "Post"), each=n)
  ) %>% 
    spread(PrePost, Outcome)
  
  out <- list()
  
  out$apparent_worse <- round(100*with(measured, mean(Post > Pre)),1)
  out$apparent_10    <- round(100*with(measured, mean((Pre-Post)/Pre > 0.1)),1)
  out$apparent_20    <- round(100*with(measured, mean((Pre-Post)/Pre > 0.2)),1)
  out$true_worse     <- round(100*with(true, mean(Post > Pre)),1)
  out$true_10        <- round(100*with(true, mean((Pre-Post)/Pre > 0.1)),1)
  out$true_20        <- round(100*with(true, mean((Pre-Post)/Pre > 0.2)),1)
  
  return(out)

}

if(!file.exists("../DerivedData/percdifs.rds") || overwrite) {

measured_percs <- tidyr::crossing(
  delta = seq(0, 3, by=0.5),
  cv_delta = c(0, 0.5),
  decomp=c(1,2)
) %>%
  mutate(condition = 1:n()) %>%
  group_by(condition) %>%
  nest() %>%
  mutate(res = map(data, ~apparent_effects( n=1e7,
                                               delta=.x$delta,
                                               cv_delta = .x$cv_delta,
                                               decomp=.x$decomp))) %>%
  ungroup()

saveRDS(measured_percs, "../DerivedData/percdifs.rds")

}
```


```{r}
measured_percs <- readRDS("../DerivedData/percdifs.rds")

measured_percs_summary <- measured_percs %>% 
  mutate(apparent_worse = map_dbl(res, "apparent_worse"),
         apparent_10 = map_dbl(res, "apparent_10"),
         apparent_20 = map_dbl(res, "apparent_20"),
         true_worse = map_dbl(res, "true_worse"),
         true_10 = map_dbl(res, "true_10"),
         true_20 = map_dbl(res, "true_20")) %>% 
  select(-res) %>% 
  unnest(data) %>% 
  #mutate(true_worse = round(100*(1-pnorm(1/cv_delta)),1)) %>% 
  arrange(decomp, delta, cv_delta) %>% 
  mutate(decomp = ifelse(decomp==1, 
                         trt_all$decomp[1], 
                         trt_all$decomp[2]),
         cv_delta = ifelse(cv_delta==0,
                           "Homogeneous",
                           "Heterogeneous")) %>% 
  select(decomp, condition, delta, cv_delta,
         true_worse, apparent_worse,
         true_10, apparent_10, 
         true_20, apparent_20) %>% 
  rename("Patients" = decomp,
         "Apparent Worsening (%)" = apparent_worse,
         "True Worsening (%)" = true_worse,
         "True 10%+ Improvement (%)" = true_10,
         "Apparent 10%+ Improvement (%)" = apparent_10,
         "True 20%+ Improvement (%)" = true_20,
         "Apparent 20%+ Improvement (%)" = apparent_20,
         "Effects" = cv_delta,
         "Difference" = delta) %>% 
  select(-condition)

decomp_change <- head(
  which(
    measured_percs_summary$Patients ==
      trt_all$decomp[2]), 1)

knitr::kable(measured_percs_summary[,-1]) %>% 
  kable_styling("striped", full_width = F) %>%
  pack_rows(trt_all$decomp[1], 1, decomp_change-1) %>%
  pack_rows(trt_all$decomp[2], decomp_change, nrow(measured_percs_summary))

appchange <- knitr::kable(measured_percs_summary[,-1]) %>% 
  kable_styling("striped", full_width = F) %>%
  pack_rows(trt_all$decomp[1], 1, decomp_change-1) %>%
  pack_rows(trt_all$decomp[2], decomp_change, nrow(measured_percs_summary))

appchange

# save_kable(appchange, file = "figures/appchange.jpg")
```

So, for a true effect of about 2mmHg in compensated patients, it will appear as if 12-16% would appear to worsen.  This fits with clinical experience.


## Simulation

Note: these are no longer run as we instead make use of the analytical solutions

```{r}
HVPG_dif_sim <- function(n, delta, cv_delta, wscv=trt_all$wscv, 
                     mean=trt_all$mean, cv=trt_all$cv, icc=trt_all$icc,
                     decomp = 1) {
  
  wscv <- wscv[decomp]
  mean <- mean[decomp]
  cv   <- cv[decomp]
  
  var <- (cv*mean)^2
  sd_true <- sqrt(var * icc)
  
  pre_true <- rnorm(n, mean, sd_true)
  
  pre_meas <- pre_true + rnorm(n, 0, mean*wscv)
  
  post_true <- pre_true - rnorm(n, delta, cv_delta*delta)
  
  post_meas <- post_true + rnorm(n, 0, mean*wscv)
  
  measured <- tibble::tibble(
    ID = rep(1:n, times=2),
    Outcome = c(pre_meas, post_meas),
    PrePost = rep(c("Pre", "Post"), each=n)
  )
  
  d <- effsize::cohen.d(measured$Outcome, measured$PrePost, 
                        paired=TRUE)$estimate
  
  dz <- effsize::cohen.d(measured$Outcome, measured$PrePost, 
                        paired=TRUE, within=FALSE)$estimate
  
  test <- t.test(pre_meas, post_meas, alternative = "greater", 
              paired = T)
  
  # Note: one-sided p value
  
  testout <- broom::tidy(test)
  
  out <- mutate(testout, d = d, dz = dz)
  
  return(out)
  
}
```


Now, we set up the simulation parameters for various scenarios.

```{r}
difsimpars <- tidyr::crossing(
  n=seq(5, 100, by = 5),
  delta = seq(1,3, by=0.5),
  cv_delta = c(0, 0.5),
)
```

And now we run them

```{r, eval=FALSE}
if(!file.exists(paste0("../DerivedData/difsims_decomp_", 
                       nsims, ".rds")) || overwrite) {
  
  pb <- progress_bar$new(total = nrow(difsimpars))
  
  difsims <- difsimpars %>% 
    mutate(sim = 1:nrow(difsimpars)) %>% 
    group_by(sim) %>% 
    nest(params = c(n, delta, cv_delta)) %>% 
    mutate(output = map(params,
                        ~{pb$tick(); 
                          bind_rows(purrr::rerun(nsims, 
                            HVPG_dif_sim(.x$n, .x$delta, 
                                       .x$cv_delta, 
                                       decomp=1)))}))

  saveRDS(difsims, paste0("../DerivedData/difsims_decomp_", nsims, ".rds"))
  
}
```

```{r, eval=FALSE}
if(!file.exists(paste0("../DerivedData/difsims_comp_", 
                       nsims, ".rds")) || overwrite) {
  
  pb <- progress_bar$new(total = nrow(difsimpars))
  
  difsims <- difsimpars %>% 
    mutate(sim = 1:nrow(difsimpars)) %>% 
    group_by(sim) %>% 
    nest(params = c(n, delta, cv_delta)) %>% 
    mutate(output = map(params,
                        ~{pb$tick(); 
                          bind_rows(purrr::rerun(nsims, 
                            HVPG_dif_sim(.x$n, .x$delta, 
                                       .x$cv_delta,  
                                       decomp=2)))}))

  saveRDS(difsims, paste0("../DerivedData/difsims_comp_", nsims, ".rds"))
  
}
```


And extract the results

```{r}
difsims_decomp <- readRDS(
  paste0("../DerivedData/difsims_decomp_", nsims, ".rds"))

difsims_decomp_res <- difsims_decomp %>% 
  ungroup() %>% 
  mutate(power = map_dbl(output, ~mean(.x$p.value < 0.05))) %>% 
  unnest(params) %>% 
  mutate(delta = as.factor(delta)) %>% 
  mutate(Effects = ifelse(cv_delta==0, "Homogeneous Effects", 
                                       "Heterogeneous Effects")) %>% 
  mutate(Effects = fct_inorder(Effects)) %>% 
  mutate(decomp = trt_all$decomp[1])





difsims_comp <- readRDS(
  paste0("../DerivedData/difsims_comp_", nsims, ".rds"))

difsims_comp_res <- difsims_comp %>% 
  ungroup() %>% 
  mutate(power = map_dbl(output, ~mean(.x$p.value < 0.05))) %>% 
  unnest(params) %>% 
  mutate(delta = as.factor(delta)) %>% 
  mutate(Effects = ifelse(cv_delta==0, "Homogeneous Effects", 
                                       "Heterogeneous Effects")) %>% 
  mutate(Effects = fct_inorder(Effects)) %>% 
  mutate(decomp = trt_all$decomp[2])




difsims_res <- bind_rows(difsims_comp_res, difsims_decomp_res)
```

## Plotting

```{r, fig.height=6, fig.width=7, message=FALSE, warning=FALSE, eval=FALSE}
ggplot(difsims_res, aes(x=n, y=power, colour=delta)) +
  geom_point() + 
  geom_line() +
  facet_grid(decomp~Effects) +
  coord_cartesian(ylim=c(0.5, 1)) +
  scale_color_brewer(type = "qual", palette = 2) +
  annotate("rect", ymin = 0.8, ymax = 1, xmin=0, 
           xmax=100, alpha = .4, fill="grey") +
  labs(y="Power", x="Sample Size", 
       colour="Intervention\nEffect (mmHg)")
```

## Required Individuals

And how many people do we need for each scenario?

```{r, eval=TRUE}
difsims_80power <- difsims_res %>% 
  arrange(power) %>% 
  filter(power > 0.8) %>% 
  select(decomp, delta, Effects, n) %>% 
  group_by(delta, Effects, decomp) %>% 
  slice(1) %>% 
  arrange(decomp, delta) %>% 
  rename("Patients" = decomp,
         "Difference (mmHg)" = delta,
         "80% Power" = n)

difsims_90power <- difsims_res %>% 
  arrange(power) %>% 
  filter(power > 0.9) %>% 
  select(decomp, delta, Effects, n) %>% 
  group_by(delta, Effects, decomp) %>% 
  slice(1) %>% 
  arrange(decomp, delta) %>% 
  rename("Patients" = decomp,
         "Difference (mmHg)" = delta,
         "90% Power" = n)

difsims_power <- left_join(difsims_80power, difsims_90power)

decomp_change <- head(
  which(
    difsims_power$Patients ==
      trt_all$decomp[2]), 1)

# kable(difsims_power[,-1]) %>% 
#   kable_styling("striped", full_width = F) %>%
#   pack_rows(trt_all$decomp[1], 1, decomp_change-1) %>%
#   pack_rows(trt_all$decomp[2], decomp_change, nrow(difsims_80power))
```


## Analytical solution


```{r}
difsims_ana <- function(Patients, Difference, Effects, Power,
                        trt_all=trt_all) {
  
  Effects <- as.character(Effects)
  Difference <- as.numeric(Difference)
  
  heterogen <- ifelse(Effects=="Homogeneous Effects",
                      0, 0.5)
  
  decomp = ifelse(Patients=="Includes Decompensated", 1, 2)
  
  trt_all_pat <- trt_all[decomp,]
  
  signvar_sd <- sqrt(
    (trt_all_pat$signvar_sd*trt_all_pat$mean)^2 +
    (heterogen*Difference)^2)
  
  dz <- Difference/signvar_sd
  
  ceiling(pwr::pwr.t.test(d=dz, sig.level = 0.05, power = Power,
                       alternative = "greater", type = "paired")$n)
  
}

difsims_power_ana <- difsims_power %>% 
  rename(Difference = `Difference (mmHg)`) %>% 
  mutate(Difference = as.numeric(as.character(Difference)),
         Effects= as.character(Effects)) %>% 
  group_by(Patients, Difference, Effects) %>% 
  mutate("80% Power"= pmap_dbl(list(Patients, Difference,
                                    Effects), difsims_ana, Power=0.8,
                                   trt_all = trt_all),
         "90% Power"= pmap_dbl(list(Patients, Difference,
                                    Effects), difsims_ana, Power=0.9,
                                   trt_all = trt_all))

kable(difsims_power_ana[,-1]) %>% 
  kable_styling("striped", full_width = F) %>%
  pack_rows(trt_all$decomp[1], 1, decomp_change-1) %>%
  pack_rows(trt_all$decomp[2], decomp_change, nrow(difsims_power_ana))

difpower <- kable(difsims_power_ana[,-1]) %>% 
  kable_styling("striped", full_width = F) %>%
  pack_rows(trt_all$decomp[1], 1, decomp_change-1) %>%
  pack_rows(trt_all$decomp[2], decomp_change, nrow(difsims_power_ana))

difpower

# save_kable(difpower, "figures/difpower.jpg")
```


### Contour Plots

```{r}
difsims_ana_contour <- function(Patients, Difference, Effects, n,
                        trt_all=trt_all) {
  
  Effects <- as.character(Effects)
  Difference <- as.numeric(Difference)
  
  heterogen <- ifelse(Effects=="Homogeneous Effects",
                      0, 0.5)
  
  decomp = ifelse(Patients=="Includes Decompensated", 1, 2)
  
  trt_all_pat <- trt_all[decomp,]
  
  signvar_sd <- sqrt(
    (trt_all_pat$signvar_sd*trt_all_pat$mean)^2 +
    (heterogen*Difference)^2)
  
  dz <- Difference/signvar_sd
  
  pwr::pwr.t.test(d=dz, sig.level = 0.05, n = n,
                       alternative = "greater", type = "paired")$power
  
}

contour_dat <- tidyr::crossing(Difference = seq(0.3,3,length.out=2701),
                               n = 5:80,
                               Effects = c("Homogeneous Effects", 
                                           "Heterogeneous Effects"),
                               Patients = c("Includes Decompensated", 
                                            "Only Compensated"))

```


Run it

```{r, eval=FALSE}
contour_power <- contour_dat %>% 
  mutate(Test = 1:n()) %>% 
  group_by(Test) %>% 
  mutate(Power=pmap_dbl(list(Patients, Difference, Effects, n), 
                        difsims_ana_contour, trt_all=trt_all)) %>% 
  ungroup()

saveRDS(contour_power, "../DerivedData/contour_power.rds")
```


```{r, fig.height=5, fig.width=10}
contour_power <- readRDS("../DerivedData/contour_power.rds")

library(viridis)

contour_power <- contour_power %>% 
  mutate(Power_cut = cut(Power, breaks = seq(0,1, length.out = 11))) %>% 
  mutate(Power = ifelse(Power == 1, 0.99, Power)) # Note below to explain this
                                                  

contour_het <- contour_power %>% 
  filter(Effects!="Homogeneous Effects")

contour_hom <- contour_power %>% 
  filter(Effects=="Homogeneous Effects")


homplot <- ggplot(contour_hom, aes(x=n, y=Difference, z=Power)) +
  geom_contour_filled(alpha=0.8, breaks=seq(0,1, by=0.1)) +
  facet_wrap(.~Patients, scales = "free") +
  theme_ipsum_rc() +
  labs(x = "Sample Size",
       y = "Hypothetical True Intervention Effect (mmHg)") +
  scale_y_continuous(breaks = seq(0.5, 3, by = 0.5)) +
  scale_fill_viridis("Power", discrete = T) +
  geom_contour(breaks=0.8, colour="black", linetype="dashed")


hetplot <- ggplot(contour_het, aes(x=n, y=Difference, z=Power)) +
  geom_contour_filled(alpha=0.8, breaks=seq(0,1.1, by=0.1)) +
  facet_wrap(.~Patients, scales = "free") +
  theme_ipsum_rc() +
  labs(x = "Sample Size",
       y = "Hypothetical True Intervention Effect (mmHg)") +
  scale_y_continuous(breaks = seq(0.5, 3, by = 0.5)) +
  scale_fill_viridis("Power", discrete = T) +
  geom_contour(breaks=0.8, colour="black", linetype="dashed")

homplot

hetplot

ggsave(homplot, height=5, width=10, filename = "figures/Dif_hom_contour.png")
ggsave(hetplot, height=5, width=10, filename = "figures/Dif_het_contour.png")

ggsave(homplot, height=5, width=10, filename = "figures/Dif_hom_contour.jpg", 
       dpi = 600)
ggsave(hetplot, height=5, width=10, filename = "figures/Dif_het_contour.jpg", 
       dpi = 600)

# +
#   xlab("Minor allele frequency")+
#   ylab("Hypothetical effect size")+
#   scale_x_log10(expand = c(0, 0), position = "bottom") +
#   scale_y_continuous(expand = c(0, 0))+
#   geom_line(data = power.80, col = "black")+
#   theme_classic(base_size = 12)
```

The strange line of code where I convert those values = 1 to 0.99 is to prevent the graph from showing strangely. I think this has to do with floating point accuracy.  Some of the values equal to 1, are rounded in the computer number system to a little bit above 1, and then the plot makes them white. So by setting them to 0.99, they're still the same colour, but the plot is filled correctly.


# Power Analysis for Difference in Differences

## Simulation

```{r}
HVPG_difindif_sim <- function(n, delta1, delta2, cv_delta, 
                              wscv=trt_all$wscv, 
                              mean=trt_all$mean, 
                              cv=trt_all$cv, 
                              decomp ) {
  
  wscv <- wscv[decomp]
  mean <- mean[decomp]
  cv   <- cv[decomp]
  
  var <- (cv*mean)^2
  sd_true <- sqrt(var * icc)
  
  pre_true1 <- rnorm(n, mean, sd_true)
  pre_true2 <- rnorm(n, mean, sd_true)
  
  pre_meas1 <- pre_true1 + rnorm(n, 0, mean*wscv)
  pre_meas2 <- pre_true2 + rnorm(n, 0, mean*wscv)
  
  post_true1 <- pre_true1 - rnorm(n, delta1, cv_delta*delta1)
  post_true2 <- pre_true2 - rnorm(n, delta2, cv_delta*delta2)
  
  post_meas1 <- post_true1 + rnorm(n, 0, mean*wscv)
  post_meas2 <- post_true2 + rnorm(n, 0, mean*wscv)
  
  measured <- tibble::tibble(
    ID = rep(1:n, times=2),
    Pre = c(pre_meas1, pre_meas2),
    Post = c(post_meas1, post_meas2),
    Diff = Post - Pre,
    Group = rep(c("A", "B"), each=n)
  )
  
  d <- effsize::cohen.d(measured$Diff, measured$Group, 
                        paired=FALSE)$estimate
  
  mod <- lm(Post ~ Pre + Group, data=measured)
  
  testout <- broom::tidy(mod) %>% 
    filter(term=="GroupB") %>% 
    select(-term) %>% 
    mutate(p.value = pt(statistic, mod$df, lower.tail = FALSE))
  # Note: using a one-sided p value
  
  out <- mutate(testout, d=d)
  
  return(out)
  
}
```

Now, we set up the simulation for various scenarios.

```{r}
difindifsimpars <- tidyr::crossing(
  n = c( seq(5, 100, by = 5), seq(110, 200, by=10)),
  delta1 = seq(1,3, by = 0.5),
  delta2 = seq(0,2, by = 0.5),
  cv_delta = c(0, 0.5),
) %>% 
  mutate(deltadif = delta1 - delta2) %>% 
  filter(delta1 > delta2)
```

And we run it

```{r}
# if(!file.exists(paste0("../DerivedData/difindifsims_decomp_", 
#                        nsims, ".rds")) || overwrite) {
#   
#   pb <- progress_bar$new(total = nrow(difindifsimpars))
#   
#   difindifsims <- difindifsimpars %>% 
#     mutate(sim = 1:nrow(difindifsimpars)) %>% 
#     group_by(sim) %>% 
#     nest(params = c(n, delta1, delta2, cv_delta)) %>% 
#     mutate(output = map(params, .progress = TRUE,
#                         ~{pb$tick(); 
#                           bind_rows(purrr::rerun(nsims, 
#                           HVPG_difindif_sim(.x$n, .x$delta1, 
#                                             .x$delta2, .x$cv_delta, 
#                                             decomp = 1)))}))
#   
#   saveRDS(difindifsims, 
#           paste0("../DerivedData/difindifsims_decomp_", nsims, ".rds"))
#   
# }
  

```





```{r, eval=FALSE}
# if(!file.exists(paste0("../DerivedData/difindifsims_comp_", 
#                        nsims, ".rds")) || overwrite) {
#   
#   pb <- progress_bar$new(total = nrow(difindifsimpars))
#   
#   difindifsims <- difindifsimpars %>% 
#     mutate(sim = 1:nrow(difindifsimpars)) %>% 
#     group_by(sim) %>% 
#     nest(params = c(n, delta1, delta2, cv_delta)) %>% 
#     mutate(output = map(params, 
#                         ~{pb$tick(); 
#                           bind_rows(purrr::rerun(nsims, 
#                           HVPG_difindif_sim(.x$n, .x$delta1, 
#                                             .x$delta2, .x$cv_delta,
#                                             decomp = 2)))}))
#   
#   saveRDS(difindifsims, 
#           paste0("../DerivedData/difindifsims_comp_", nsims, ".rds"))
#   
# }
```

```{r, fig.height=8, fig.width=4}
# difindifsims_decomp <- readRDS(paste0("../DerivedData/difindifsims_decomp_", nsims, ".rds"))
# 
# difindifsims_decomp_res <- difindifsims_decomp %>% 
#   ungroup() %>% 
#   mutate(power = map_dbl(output, ~mean(.x$p.value < 0.05))) %>% 
#   unnest(params) %>% 
#   mutate(deltadif = delta1 - delta2,
#          delta1 = as.factor(delta1),
#          delta2 = as.factor(delta2),
#          deltadif = as.factor(deltadif)) %>% 
#   #filter(deltadif != 0.5) %>% 
#   mutate(Effects = ifelse(cv_delta==0, "Homogeneous Effects", 
#                                        "Heterogeneous Effects")) %>% 
#   mutate(Effects = fct_inorder(Effects)) %>% 
#   mutate(decomp = trt_all$decomp[1])
# 
# 
# difindifsims_comp <- readRDS(paste0("../DerivedData/difindifsims_comp_", nsims, ".rds"))
# 
# difindifsims_comp_res <- difindifsims_comp %>% 
#   ungroup() %>% 
#   mutate(power = map_dbl(output, ~mean(.x$p.value < 0.05))) %>% 
#   unnest(params) %>% 
#   mutate(deltadif = delta1 - delta2,
#          delta1 = as.factor(delta1),
#          delta2 = as.factor(delta2),
#          deltadif = as.factor(deltadif)) %>% 
#   #filter(deltadif != 0.5) %>% 
#   mutate(Effects = ifelse(cv_delta==0, "Homogeneous Effects", 
#                                        "Heterogeneous Effects")) %>% 
#   mutate(Effects = fct_inorder(Effects)) %>% 
#   mutate(decomp = trt_all$decomp[2])
# 
# 
# 
# difindifsims_res <- bind_rows(difindifsims_decomp_res, 
#                               difindifsims_comp_res)

```

Now run in the cluster, and extract the results

```{r, eval=FALSE}
difindifsimfiles <- list.files("../Cluster/", pattern = "\\d.rds", 
                               full.names = T)

a <- as.numeric(str_match(difindifsimfiles, pattern = "_comp_(\\d*).rds")[,2])
which(!(1:500 %in% a))

b <- as.numeric(str_match(difindifsimfiles, pattern = "_decomp_(\\d*).rds")[,2])
which(!(1:500 %in% b))


difindifsims <- tibble(file = difindifsimfiles) %>% 
  mutate(decomp = ifelse(str_detect(file, "_decomp_"), 
                          "Includes Decompensated", 
                          "Only Compensated"),
         batch = as.numeric(str_match(file, "(\\d+).rds")[,2]),
         sim_add = 50*(batch-1)) %>% 
  rowwise() %>% 
  mutate(data = map(file, readRDS))

difindifsims <- difindifsims %>% 
  ungroup() %>% 
  unnest(data) %>% 
  ungroup() %>% 
  unnest(params) %>% 
  unnest(output)

difindifsims_res <- difindifsims %>% 
  group_by(decomp, deltadif, n, delta1, delta2, cv_delta) %>% 
  summarise(
    power = mean(p.value < 0.05),
    d = mean(d)
  ) %>% 
  ungroup() %>% 
  mutate(Effects = ifelse(cv_delta==0, "Homogeneous Effects", 
                                        "Heterogeneous Effects")) %>% 
  mutate(delta1 = as.factor(delta1),
         delta2 = as.factor(delta2),
         deltadif = as.factor(deltadif))

saveRDS(difindifsims_res, "../DerivedData/difindifsims_res.rds")
```
```{r}
difindifsims_res <- readRDS("../DerivedData/difindifsims_res.rds")

difindifsims_decomp_res <- filter(difindifsims_res, 
                                  decomp=="Includes Decompensated")

difindifsims_comp_res <- filter(difindifsims_res, 
                                  decomp=="Only Compensated")
```



## Plotting

Here we have the size of the effect of the better intervention as the 

```{r, fig.height=7, fig.width=11, message=FALSE, warning=FALSE}
difindifplot_decomp <-
  ggplot(difindifsims_decomp_res, aes(x=n, y=power, colour=delta2)) +
    geom_point() + 
    geom_line() +
    facet_grid(delta1~Effects) +
    coord_cartesian(ylim=c(0.5, 1)) +
    annotate("rect", ymin = 0.8, ymax = 1, xmin=0, xmax=200, 
             alpha = .4, fill="grey") +
    scale_color_brewer(type = "qual", palette = 2) +
    labs(y="Power", x="Sample Size", 
       colour="Reference\nIntervention\nEffect (mmHg)",
       title=trt_all$decomp[1]) +
    theme(plot.title = element_text(hjust = 0.5))

difindif_legend <- cowplot::get_legend(difindifplot_decomp)

difindifplot_decomp <- difindifplot_decomp +
  guides(colour=FALSE)

difindifplot_comp <-
  ggplot(difindifsims_comp_res, aes(x=n, y=power, colour=delta2)) +
    geom_point() + 
    geom_line() +
    facet_grid(delta1~Effects) +
    coord_cartesian(ylim=c(0.5, 1)) +
    annotate("rect", ymin = 0.8, ymax = 1, xmin=0, xmax=200, 
             alpha = .4, fill="grey") +
    scale_color_brewer(type = "qual", palette = 2) +
    labs(y="Power", x="Sample Size", 
       colour="Reference\nIntervention\nEffect (mmHg)",
       title=trt_all$decomp[2]) +
    theme(plot.title = element_text(hjust = 0.5)) +
    guides(colour=FALSE)

difindifplot <- cowplot::plot_grid(
  difindifplot_decomp, difindifplot_comp, difindif_legend,
  nrow = 1, rel_widths = c(3,3,0.5)
)

difindifplot
```

## Required Individuals

And how many people do we need for each scenario?

```{r}
difindifsims_80power <- difindifsims_res %>% 
  arrange(power) %>% 
  filter(power > 0.8) %>% 
  select(decomp, deltadif, delta1, delta2, Effects, n) %>% 
  group_by(delta1, deltadif, Effects, decomp) %>% 
  slice(1) %>% 
  arrange(decomp, desc(deltadif), desc(delta1), Effects) %>% 
  rename("Patients" = decomp,
         "Intervention 1 Effect (mmHg)" = delta1,
         "Intervention 2 Effect (mmHg)" = delta2,
         "Difference (mmHg)" = deltadif,
         "80% Power" = n)

difindifsims_90power <- difindifsims_res %>% 
  arrange(power) %>% 
  filter(power > 0.9) %>% 
  select(decomp, deltadif, delta1, delta2, Effects, n) %>% 
  group_by(delta1, deltadif, Effects, decomp) %>% 
  slice(1) %>% 
  arrange(decomp, desc(deltadif), desc(delta1), Effects) %>% 
  rename("Patients" = decomp,
         "Intervention 1 Effect (mmHg)" = delta1,
         "Intervention 2 Effect (mmHg)" = delta2,
         "Difference (mmHg)" = deltadif,
         "90% Power" = n)

difindifsims_power <- left_join(difindifsims_80power, difindifsims_90power) %>%
  mutate(`90% Power` = as.character(`90% Power`)) %>% 
  mutate(`90% Power` = ifelse(is.na(`90% Power`),
                              ">200", `90% Power`)) %>% 
  filter(`Difference (mmHg)` != 0.5)

decomp_change <- head(
  which(
    difindifsims_power$Patients ==
      trt_all$decomp[2]), 1)

kable(difindifsims_power[,-1]) %>% 
  kable_styling("striped", full_width = F) %>%
  pack_rows(trt_all$decomp[1], 1, decomp_change-1) %>%
  pack_rows(trt_all$decomp[2], decomp_change, nrow(difindifsims_power))

difindiftable_decomp <- difindifsims_power %>% 
  filter(Patients=="Includes Decompensated") %>% 
  ungroup() %>% 
  select(-Patients) %>% 
  kable() %>% 
  kable_styling("striped", full_width = F) %>%
  pack_rows("Includes Decompensated", 1, decomp_change-1) 

difindiftable_comp <- difindifsims_power %>% 
  filter(Patients=="Only Compensated") %>% 
  ungroup() %>% 
  select(-Patients) %>% 
  kable() %>% 
  kable_styling("striped", full_width = F) %>%
  pack_rows("Only Compensated", 1, decomp_change-1)

difindiftable_decomp
difindiftable_comp

# save_kable(difindiftable_decomp, "figures/difindifpower_dc.jpg")
# save_kable(difindiftable_comp, "figures/difindifpower_c.jpg")
```



## Contour Plot

First let's look at different deltadif values.

```{r, fig.height=15, fig.width=10}
difindifsims_res_hom <- difindifsims_res %>% 
  filter(cv_delta==0) %>% 
  mutate(delta1 = as.numeric(as.character(delta1)),
         delta2 = as.numeric(as.character(delta2)),
         deltadif = as.numeric(as.character(deltadif))) %>% 
  mutate(power = ifelse(power == 1, 0.99, power)) 

difindifsims_res_het <- difindifsims_res %>% 
  filter(cv_delta!=0) %>% 
  mutate(delta1 = as.numeric(as.character(delta1)),
         delta2 = as.numeric(as.character(delta2)),
         deltadif = as.numeric(as.character(deltadif))) %>% 
  mutate(power = ifelse(power == 1, 0.99, power)) 

difindif_cont1_hom <- difindifsims_res_hom %>% 
  filter(deltadif!=3) %>% 
  ggplot(aes(x=n, y=delta1, z=power)) +
  geom_contour_filled(alpha=0.8, breaks=seq(0,1.1, by=0.1)) +
  facet_grid(deltadif~decomp, scales = "free") +
  theme_ipsum_rc() +
  labs(x = "Sample Size",
       y = "Hypothetical Intervention Effect of Main Treatment (mmHg)",
       subtitle = "Comparison of Differences of Effects between Interventions") +
  scale_y_continuous(breaks = seq(0.5, 5, by = 0.5)) +
  scale_fill_viridis("Power", discrete = T) +
  geom_contour(breaks=0.8, colour="black", linetype="dashed")

difindif_cont1_hom

difindif_cont1_het <- difindifsims_res_het %>% 
  filter(deltadif!=3) %>% 
  ggplot(aes(x=n, y=delta1, z=power)) +
  geom_contour_filled(alpha=0.8, breaks=seq(0,1.1, by=0.1)) +
  facet_grid(deltadif~decomp, scales = "free") +
  theme_ipsum_rc() +
  labs(x = "Sample Size",
       y = "Hypothetical Intervention Effect of Main Treatment (mmHg)",
       subtitle = "Comparison of Differences of Effects between Interventions") +
  scale_y_continuous(breaks = seq(0.5, 5, by = 0.5)) +
  scale_fill_viridis("Power", discrete = T) +
  geom_contour(breaks=0.8, colour="black", linetype="dashed")    

difindif_cont1_het
```
This is helpful to visualise, though probably not for the paper. With homogeneous effects, the determinant of the power is the difference in intervention effects; it's a straight line. With heterogeneous effects, the line is not straight


```{r, fig.height=15, fig.width=10}
difindif_cont2_hom <- ggplot(difindifsims_res_hom, aes(x=n, y=delta1, z=power)) +
  geom_contour_filled(alpha=0.8, breaks=seq(0,1.1, by=0.1)) +
  facet_grid(delta2~decomp, scales = "free") +
  theme_ipsum_rc() +
  labs(x = "Sample Size",
       y = "Hypothetical Intervention Effect of Main Treatment (mmHg)",
       subtitle = "Comparisons Grouped by Intervention Effects of the Reference Treatment (mmHg)") +
  scale_y_continuous(breaks = seq(0.5, 5, by = 0.5)) +
  scale_fill_viridis("Power", discrete = T) +
  geom_contour(breaks=0.8, colour="black", linetype="dashed")

difindif_cont2_hom

difindif_cont2_het <- ggplot(difindifsims_res_het, aes(x=n, y=delta1, z=power)) +
  geom_contour_filled(alpha=0.8, breaks=seq(0,1.1, by=0.1)) +
  facet_grid(delta2~decomp, scales = "free") +
  theme_ipsum_rc() +
  labs(x = "Sample Size",
       y = "Hypothetical Intervention Effect of Main Treatment (mmHg)",
       subtitle = "Comparisons Grouped by Intervention Effects of the Reference Treatment (mmHg)") +
  scale_y_continuous(breaks = seq(0.5, 5, by = 0.5)) +
  scale_fill_viridis("Power", discrete = T) +
  geom_contour(breaks=0.8, colour="black", linetype="dashed")    

difindif_cont2_het
```

And just looking against placebo (i.e. delta2=0)

```{r, fig.height=5, fig.width=10}
difindif_cont3_hom <- difindifsims_res_hom %>% 
  filter(deltadif!=3) %>% 
  filter(delta2==0) %>% 
  ggplot(aes(x=n, y=delta1, z=power)) +
  geom_contour_filled(alpha=0.8, breaks=seq(0,1.1, by=0.1)) +
  facet_wrap(.~decomp, scales = "free") +
  theme_ipsum_rc() +
  labs(x = "Sample Size",
       y = "Hypothetical Intervention Effect of Main Treatment (mmHg)") +
  scale_y_continuous(breaks = seq(0.5, 5, by = 0.5)) +
  scale_fill_viridis("Power", discrete = T) +
  geom_contour(breaks=0.8, colour="black", linetype="dashed") +
  xlim(c(5,120))

difindif_cont3_hom

difindif_cont3_het <- difindifsims_res_het %>% 
  filter(deltadif!=3) %>% 
  filter(delta2==0) %>% 
  ggplot(aes(x=n, y=delta1, z=power)) +
  geom_contour_filled(alpha=0.8, breaks=seq(0,1.1, by=0.1)) +
  facet_wrap(.~decomp, scales = "free") +
  theme_ipsum_rc() +
  labs(x = "Sample Size",
       y = "Hypothetical Intervention Effect of Main Treatment (mmHg)") +
  scale_y_continuous(breaks = seq(0.5, 5, by = 0.5)) +
  scale_fill_viridis("Power", discrete = T) +
  geom_contour(breaks=0.8, colour="black", linetype="dashed") +
  xlim(c(5,120))

difindif_cont3_het

ggsave(difindif_cont3_hom, height=5, width=10, filename = "figures/Difindif_hom_contour.png")
ggsave(difindif_cont3_het, height=5, width=10, filename = "figures/Difindif_het_contour.png")

ggsave(difindif_cont3_hom, height=5, width=10, filename = "figures/Difindif_hom_contour.jpg", 
       dpi = 600)
ggsave(difindif_cont3_het, height=5, width=10, filename = "figures/Difindif_het_contour.jpg", 
       dpi = 600)
```



# ICC Figures

```{r icc_distribs, fig.height=8, fig.width=7}
colours <- c("#61b096", "#bd7969")

make_distributions <- function(icc) {
  
  sd_true <- 1
  var_true <- sd_true^2
  
  var_tot <- var_true / icc
  sd_tot <- sqrt(var_tot)
  
  sd_error <- sqrt(var_tot - var_true)
  
  distrib <- ggplot(data.frame(x=c(-5,5)), aes(x=x)) +
      stat_function(fun = dnorm, 
        colour = "black", size = 1.5, args = list(mean = 0, sd=sd_tot), 
        geom = "line") +
    stat_function(fun = dnorm, 
        colour = colours[1], size = 1, args = list(mean = 0, sd=sd_true), 
        geom = "line") +
    stat_function(fun = dnorm, 
        colour = colours[2], size = 1, args = list(mean = 0, sd=sd_error), 
        geom = "line") +
    geom_vline(xintercept = 0, linetype="dashed") +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
          axis.text.x=element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank()) +
    annotate("text", x=2.5, y=0.5, 
           label="Error\nVariance",
           colour = colours[2], hjust=0.5) +
    annotate("text", x=-2.5, y=0.5, 
           label="True\nVariance",
           colour = colours[1], hjust=0.5) +
    coord_cartesian(ylim=c(0,0.7))
  
  return(distrib)
}

make_piecharts <- function(icc) {
  
  sd_true <- 1
  var_true <- sd_true^2
  
  var_tot <- var_true / icc
  sd_tot <- sqrt(var_tot)
  
  var_error <- var_tot - var_true
  
  var_pie <- data.frame(Variance = c("True", "Error"),
                Values = c(var_true, var_error))
  var_pie$Variance <- forcats::fct_inorder(var_pie$Variance)
  
  
  pie <- ggplot(var_pie, aes(x="", y=Values, fill=Variance))+
    geom_bar(width = 1, stat = "identity") +
    theme(axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank()) +
    coord_polar("y", start=0) +
    scale_fill_manual(values = colours) +
    labs(x="", y="") +
    guides(fill=FALSE) +
    #labs(title=paste0("ICC=", icc)) +
    NULL
    
  
  return(pie)
}

make_distrib_and_pie <- function(icc) {
  
  dist <- make_distributions(icc)
  pie <- make_piecharts(icc)
  
  outplot <- cowplot::plot_grid(pie, dist, rel_widths = c(1,2)) +
    draw_figure_label(paste0("ICC=", icc))
  
  return(outplot)
  
}

comparison_charts <- tibble(
  icc = c(0.1, 0.3, 0.7, 0.9)
) %>% 
  mutate(chart = purrr::map(icc, make_distrib_and_pie))

plot_grid(plotlist = comparison_charts$chart, ncol=1)
```

